\textsf{Are large language models superhuman chemists?}

### 2024-10-25 src/tex/appendix.tex
add table for models struggle to answer

\subsubsection{Curation workflow}
Questions were added via pull requests to the \chembench repository on GitHub.
This allowed for a manual review of each question by expert reviewers (with backgrounds in chemistry, materials science, chemical engineering, and physics).
The reviews were conducted directly on the GitHub platform, where our entire curation history is also publicly available.

The general guidelines followed by the reviewer are the following:

\begin{itemize}
    \item \textbf{Originality:} Questions should not be readily findable online or in other easily accessible sources (example \url{https://github.com/lamalab-org/chem-bench/pull/392#discussion_r1694299474})
    \item \textbf{Ambiguity:} Questions with unclear wording or multiple interpretations (example \url{https://github.com/lamalab-org/chem-bench/pull/420#discussion_r1698147159} )
    \item \textbf{Factual or heuristic Errors: }Questions containing factual inaccuracies or misconceptions are not included (example \url{https://github.com/lamalab-org/chem-bench/pull/389#discussion_r1686187301})
    \item \textbf{Out of Scope:} Questions outside the realm of chemistry are rejected.
    \item \textbf{Clarity and Difficulty: } They should pose a challenge and encourage exploration within the chemical domain.
    \item \textbf{Contribution to Dataset Diversity: }Questions should cover a wide range of chemical concepts and sub-disciplines. They should add value by expanding the breadth of the dataset. That is, questions already multiple (>10) times in the corpus in a similar form are rejected.
\end{itemize}

Reviewers also solved the questions to verify the answers. They also performed web searches to ensure questions were not easily found online. The reviewers often guided the revision process to ensure the question aligned with the guidelines. Questions that don't meet the criteria are either rejected or suggested for revision and most often, they are modified to a new question. Reviewers also provide feedback on the skill and difficulty annotations.

In addition to the manual review, we also performed automated checks to ensure the quality of the questions. The schemas, Latex templating, and other formatting aspects are checked automatically using GitHub Actions.

\subsubsection{Composition}

@@

@@

\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
        \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\

@@

\end{xltabular}

@@

\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\

        \toprule
        Required Skills & Number of Questions & Example Question \\
        \midrule
        Knowledge & 919 & Which of the following salts is an acidic salt? \\
        & & A. \ce{NH4Cl} \\
        & & B. \ce{Na2CO3} \\
        & & C. \ce{NaH2PO4} \\
        & & D. \ce{Zn(OH)Cl} \\
        \midrule
        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
        & & A. By atomic force microscopy \\
        & & B. By transform infrared spectroscopy \\
        & & C. By measuring a photoluminescence spectrum \\
        & & D. By absorption spectroscopy \\
        \midrule
        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
        \midrule
        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
        \bottomrule

    \label{tab:chembench_corpus_cognitive}
\end{xltabular}

@@

\clearpage
\subsection{Model performance} \label{sec:model_performance_app}
We also evaluated the model performance on the entire \chembench corpus.
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models.

@@

\Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus. The performance is measured as the fraction of questions answered completely correctly by the models.
    A score of 1 (full coverage until the outer line of this plot) indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
    }
    \label{fig:all_questions_models_completely_correct_radar_overall}
    \script{analyze_model_reports.py}
\end{figure}

@@

\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.

\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset. The relative performance difference between both is quite similar across most models. 
This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.

An interesting observation is the significant impact of chemical preference tasks on \GPTFour's scores. A detailed breakdown of overall accuracy into scores on different skills and difficulty levels is provided in \Cref{table:performance_table} and \Cref{tab:performance_table_human_subset} for \chembench corpus and the \chembenchmini subset, respectively.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic_tiny.pdf}
    \caption{\textbf{Fraction of  correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic_tiny}
    \script{analyze_performance_per_source.py}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics{figures/corpus_human_comparison.pdf}
    \caption{\textbf{
        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both the \chembench corpus and the \chembenchmini subset.
        We see that the model ranking remains fairly consistent across both sets}  
    \label{fig:performance_corpus_and_tiny}
    \script{corpus_humanset_performance.py}
\end{figure}

\begin{table}
    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
    \label{tab:performance_table_human_subset}
\end{table}

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals irregardless of molecular complexity). For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} }
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
\end{figure}

@@

\clearpage
\subsection{Influence of model scale}
\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
We see that the performance of the models correlates with their size for the models of the LLama-3 and Llama-3.1 herd of models.

@@

In our parsing workflow, we use pipelines based on regular expressions to extract the answers. In some cases, however, the answers are not directly extractable from the responses, for instance, when the model does not follow the formatting instructions. In these cases, we use a fallback mechanism to extract the answers. The fallback mechanism uses an \gls{llm} to extract the answers from the responses. The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer. We used \LlamaThreeSeventyBInstruct, accessed via the Groq API.
\Cref{tab:refusal_counts_and_parsing} tabulates the number of times the fallback mechanism was used for each model.

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one. A score of one would mean that a volunteer answered all questions correctly. A score of zero would mean that no question was answered correctly.
    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.
    }
    \label{fig:human_score_distribution}
\end{figure}

>>>>

@@

%\Cref{fig:flesch_kincaid_reading_ease} shows the distribution of the Flesch-Kincaid reading ease scores of the questions.
%We see that the questions are generally complex to read.

% \begin{figure}[htb]
%     \centering
%     \includegraphics{figures/flesch_kincaid_reading_ease.pdf}
%     \caption{\textbf{Distribution of Flesch-Kincaid reading ease scores of the questions.} The Flesch-Kincaid reading ease score\autocite{flesch1948new} measures how easy a text is to read. It is calculated based on the average number of syllables per word and words per sentence. The higher the score, the easier the text is to read. The distribution of the questions' scores is shown in the histogram. }
%     \label{fig:flesch_kincaid_reading_ease}
%     \script{wordcloud.py}
% \end{figure}

% \begin{figure}[htb]
%     \centering
%     \includegraphics{figures/question_count_barplot_mcq_vs_general.pdf}
%     \caption{\textbf{Number of multiple choice questions vs.\ open-ended questions per topic.} The bar plot shows the number of \gls{mcq} and general questions per topic.}
%     \label{fig:question_count_barplot_mcq_vs_general}
%     \script{plot_statistics.py}
% \end{figure}

@@

\begin{table}
    \centering
    \small
    \caption{\textbf{Examples for each of the topics considered in the evaluation of the \chembench corpus.} The table shows the percentage of questions in the corpus that belong to each topic, as well as example questions.} \\
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}

@@

        \end{tabularx}
    }
\end{table}

@@

\begin{table}
    \centering
    \small
    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question. Note that the total count in this table is bigger than the \chembench corpus. This is because the same question can be annotated with two different \enquote{requires} skills, i.e. Reasoning and Calculation.} \\
    \label{tab:chembench_corpus_cognitive}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \toprule
            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
            \midrule
            Which of the following salts is an acidic salt? \\
            A. \ce{NH4Cl} \\
            B. \ce{Na2CO3} \\
            C. \ce{NaH2PO4} \\
            D. \ce{Zn(OH)Cl} \\
            \midrule
            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
            \midrule
            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
            A. By atomic force microscopy \\
            B. By transform infrared spectroscopy \\
            C. By measuring a photoluminescence spectrum \\
            D. By absorption spectroscopy \\
            \midrule
            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
            \midrule
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
            \midrule
            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
            \midrule
            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
            \bottomrule
        \end{tabularx}
    }
\end{table}

@@

In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers. The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement

\begin{table}
    \centering
    \small
    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
    \label{tab:chembench_corpus_models_failed_topic}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
            A. DESI-mass spectrometry \\
            B. XPS \\
            C. atomic force microscopy \\
            D. scanning tunneling microscopy \\
            E. secondary ion mass spectrometry \\
            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
            A. \ce{SCl6} does not exist because of steric effects \\
            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
            A. All dimension of the material must be smaller than exciton Bohr radius \\
            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
            C. The movement od charge carriers is limited in all spatial dimensions \\
            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
            What makes acrylates more reactive than typical hydrocarbons? \\
            A. The carbonyl group \\
            B. The double bond \\
            C. The hydroxyl group \\
            D. The vinyl group \\
            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
            Why are polar molecules more soluble in water than non-polar molecules? \\
            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
            B. The entropy of the system increases when non-polar molecules dissolve in water \\
            C. Water has a dipole and can easier isolate polar molecules from each other \\
            D. Water makes stable dipole-dipole interactions with polar molecules \\
            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
            A. Broader product diameter distribution than pellets or extrudates \\
            B. Facilitates reactor upscaling \\
            C. It is also called snow-ball \\
            D. Not suitable to prepare egg-shell catalysts \\
            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
            What determines the flammability of polyols? \\
            A. Their chemical structure \\
            B. Their low volatility \\
            C. Their odor threshold \\
            D. Their reaction with acids \\
        \end{tabularx}
    }
\end{table}

To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them. The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
\begin{table}
    \centering
    \small
    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
    \label{tab:chembench_corpus_models_failed_cognitive}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \toprule
            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
            \midrule
            Which of the following are true for phosphoric acid? \\
            A. Is a weak acid, so it is not toxic \\
            B. Should never be tasted \\
            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
            \midrule
            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
            \midrule
            Which of the following elements has the largest atomic radius? \\
            A. \ce{Fr} \\
            B. \ce{He} \\
            C. \ce{Li} \\
            D. \ce{Rb} \\
            \midrule
            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
            \midrule
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
            \midrule
            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
            \midrule
            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}. What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V? \\
            \bottomrule
        \end{tabularx}
    }
\end{table}

\clearpage
\subsection{Model performance}
We also evaluated the model performance on the entire \chembench corpus.
\Cref{fig:barplot_all_correct_all_questions} shows the fraction of questions that were answered correctly by the models.

@@

\Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.

\begin{table}
    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
    \label{tab:performance_table_human_subset}
\end{table}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/all_questions_models_completely_correct_radar_overall.pdf}
    \caption{\textbf{Performance of the models on the different topics of the \chembench corpus.} The radar plot shows the performance of the models on the different topics of the \chembench corpus. The performance is measured as the fraction of questions answered completely correctly by the models.
    A score of 1 indicates that all questions were answered completely correctly, while a score of 0 indicates that none were answered completely correctly.
    }
    \label{fig:all_questions_models_completely_correct_radar_overall}
    \script{analyze_model_reports.py}
\end{figure}

@@

\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic_tiny.pdf}
    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic_tiny}
    \script{analyze_performance_per_source.py}
\end{figure}

% In addition, we analyzed the performance on questions which requires calculation. For this, we manually labeled questions that require multiple calculation steps.
% We find that the ranking of models is different for questions with our without calculations (\Cref{fig:requires_cal}).

% \begin{figure}
%     \centering
%     \includegraphics{figures/model_overall_cal.pdf}
%     \caption{\textbf{Overall model performance for questions with and without calculation steps.} We find that the ranking of models changes if we evaluate them on questions that require and do not require calculations, respectively.}
%     \script{requires_calculation_plot.py}
%     \label{fig:requires_cal}
% \end{figure}

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules. For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_num_atoms.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the number of atoms.} The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals).}
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_num_atoms}
\end{figure}

@@

% \begin{table}
%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation coefficient $\rho$ for the correlation of the performance of the models with the number of atoms and the complexity of the molecules. }
%     \begin{tabularx}{\textwidth}{XXXX XXXX}
%         \toprule
%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\
%         \midrule
%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
%         \midrule
%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
%         \midrule
%         \gls{smiles} \gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
%         \bottomrule
%     \end{tabularx}
%     \label{tab:correlation_coefficients}
% \end{table}

\clearpage
\subsection{Influence of model scale}
\Cref{fig:model_size_plot} shows the performance of the models as a function of the number of parameters in the model.
We see that the performance of the models correlates with scale for the models of the LLama-3 and Llama-3.1 herd of models.

@@

In our parsing workflow, we use pipelines based on regular expressions to extract the answers. In some cases, however, the answers are not directly extractable from the responses, for instance, when the model does not follow the formatting instructions. In these cases, we use a fallback mechanism to extract the answers. The fallback mechanism uses an \gls{llm} to extract the answers from the responses. The \gls{llm} is provided with the response and the question and is prompted to only extract but not generate the answer. We used \LlamaThreeSeventyBInstruct accessed via the Groq API.
\Cref{tab:refusal_counts_and_parsing} tabulates the number of times the fallback mechanism was used for each model.

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered completely correctly.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one.}
    \label{fig:human_score_distribution}
\end{figure}
